\chapter{Corrections}
 
\newcommand{\resp}[2]{{\bf $>$ #1 (\hyperref[#2]{page \pageref{#2}})}}
\newcommand{\corr}[1]{{\bf[#1]\label{#1_correction}}}

\begin{enumerate}

\section{Chapter 1}

\item{\corr{Hypothesis} Please include whether there are overarching hypotheses guiding the
proposed research.

\resp{A short discussion of overarching questions has been added to
  the introduction.}{Hypothesis} }

\item{\corr{Alignment} Please include a discussion of statistical
  approaches to alignment that might also be gainfully used to address
  the issue of misalignment in selection analyses. You mentioned the
  use of FSA and BaliPhy during the viva; can you comment why such
  approaches were not used (e.g.\ for computational reasons)?

\resp{A discussion of misalignment and statistical alignment in
  selection analyses has been added as a new section to the
  introduction.}{Alignment} }

\section{Chapter 2}
\item{\corr{Pipeline} Please describe more regarding the technical details of the analytical
pipeline; how it was implemented (language etc.), how it integrates
into ENSEMBL, how jobs were distributed etc. As the construction of
this pipeline is the main methodological contribution of this thesis,
it is important to include details to demonstrate your role.

\resp{A subsection has been added which provides some details on the
  framework used for simulations and analysis throughout the
  thesis.}{Pipeline}
}

\item{\corr{Power} Please discuss why SLR had better power as (under a correctly
specified model) one would expect that adding distributional
assumptions, one would get a better power using the 'random effects'
approaches as implemented in PAML.

\resp{Further discussion has been added to the section describing the
  results of the simulation comparing the performance of SLR and
  PAML.}{Power}
}

\item{\corr{Overalignment} In general, how do you tell if an alignment is 'wrong'? Please also
clarify what 'overalignment' is.

\resp{Some elaboration on these details has been added to the
  Conclusions section.}{Overalignment}
}

\item{\corr{False Positives} Please explain in more detail why there is an increase in false
positives at intermediate levels of divergence in Figure 2.3.

\resp{The paragraph describing this artifact has been revised and expanded.}{False Positives}
}

\item{\corr{Applicability of Simulations} Please expand the discussion to consider how the results presented
here on mammalian genes (e.g.\ choice of aligner) may/may not be
applicable to other taxa such as viruses, expanding the comparison
with Privman et al.\ [2011].

\resp{Additional discussion has been added to the Conclusion section.}{Applicability of Simulations}
}

\section{Chapter 3}

\item{\corr{Large Trees} Why are there are significant number of trees with $>100$ sequences in
Figure 3.5?

\resp{A short explanation has been added to the text.}{Large Trees}
}

\section{Chapter 4}

\item{\corr{Synonymous Rate Variation} Sitewise variation in
  synonymous substitution rates is ubiquitous (see e.g. Kosakovsky
  Pond and Muse 2005, http://mbe.oxfordjournals.org/content/22/12/2375.full), yet all the
  methods employed assume no synonymous variation. Sites with
  unusually high synonymous rates (but with $\omega=1$) might be
  misclassified as under positive selection. To what extent are the
  results biased due to a failure to incorporate synonymous variation?
  Ideally, please present some goodness-of-fit tests to demonstrate
  that this is not a serious problem.

\resp{Two additional analyses have been performed to assess the impact
  of synonymous rate variation on the analyses presented: a
  small-scale analysis in the $\beta$-globin gene, directly comparing
  SLR, PAML and HyPhy's two-rates models, and a bulk analysis across
  the genome-wide dataset which used inferred substitutions to assess
  the extent to which synonymous rate variation may have affected the
  results. Overall, SLR does not show a significant bias towards
  inferring positive selection at sites with high synonymous rates of
  substitution; only a small fraction of identified
  positively-selected sites are likely false positives due to
  synonymous rate variation.}{Synonymous Rate Variation}}

\item{\corr{Duplicated Genes} Page 84; isn't the study of adaptive evolution in duplicated genes of
great biological interest?

\resp{Yes -- it is an area of widespread research interest and
  biological significance, but gene duplications strongly impact the
  nature of the data under analysis (i.e., altering the phylogenetic
  tree which describes a gene family's evolution) in a way that is
  unique for each duplicated family. As a result, it was beyond the
  scope of this thesis to specifically analyze th evolution of
  duplicated genes. A few additional sentences were added to the
  relevant section to highlight this point.}{Duplicated Genes}}

\item{\corr{Window Size} Please discuss the rationale for using 15 codon, nonoverlapping
windows to examine potential misalignments.

\resp{A paragraph discussing the issue of window size was added to the text.}{Window Size}}

\item{\corr{Impact of Filtering} Page 97. There is quite a strong impact of filtering on the results;
which are to be believed?

\resp{A paragraph discussing further the variability due to filtering was added to the text.}{Impact of Filtering}}

\item{\corr{Terminal Versus Internal Branches} Comparison of patterns
  of substitution between terminal and internal branches of a
  phylogeny was used as the basis for a test of adaptive evolution by
  Kosakovsky Pond et
  al.\ (http://www.ploscompbiol.org/article/info:doi/10.1371/journal.pcbi.0020062);
  are there any such patterns here?

\resp{While it would be interesting to compare patterns of evolution
  in terminal and internal branches, there would be no compelling
  biological reason to expect that any mammalian population evolved
  differently during the period of time represented by a terminal
  branch in the phylogeny under analysis compared to the time
  represented by an internal branch. Put differently, the same period
  of evolutionary time may exist as a terminal or internal branch
  depending on the density of the phylogenetic tree: in a human-mouse
  phylogeny, the entire 60 million years of rodent evolution is one
  very long terminal branch, but in the 38-species mammalian tree used
  in this thesis, only the mouse-specific portion of the same
  evolutionary period is a terminal branch.

  Furthermore, the mammalian analysis focused solely on
  non-lineage-specific adaptive evolution. The comparison of internal
  versus terminal branches in the clustered substitution analysis
  (e.g. Figure 4.5) was anecdotal and served only to highlight the
  fact that terminal branches (which were likely to contain potential
  sequencing or annotation errors) showed more high-substitution
  windows than internal branches (which were generally not prone to
  such errors).}{Terminal Versus Internal Branches}}

\section{Chapter 5}

\item{\corr{P-values and Effect Sizes} Please discuss the advantages
  and disadvantages of using sitewise p values rather than the effect
  sizes ($\omega$ and associated standard errors) when conducting
  genomic level analyses. You may also want to take a look at this
  paper (http://mbe.oxfordjournals.org/content/27/3/520.abstract)
  which describes an approach which uses the pattern of nonsynonymous
  and synonymous rate variation as the basis of e.g.\ clustering. For
  the corrections, and given that all the estimates of $\omega$ and
  associated standard errors have already been produced, do the
  results of a standard meta-analysis type approach based on effect
  size differ from the p value based approach presented here?

\resp{A subsection describing the differences between ``p-value
  combination'' methods and meta-analysis methods was added to the
  text, and a standard meta-analysis approach was implemented and
  compared to the results using the p-values alone.}{P-values and Effect Sizes}
}

\item{\corr{Gene Annotation Bias} In table 5.2, there is a lot of redundancy in the gene lists
(e.g.\ SAMHD1, TLR1/4), partly due to the multiple GO terms associated
with these genes, which are mostly immune related. To what extent do
the same genes crop up again due to the large number of terms (as
opposed to their actual importance)?

\resp{The ``Top 5 Genes'' column in that table was included mostly to
  highlight the genes with the lowest gene-wide p-values that
  contributed to the enrichment of that given term. While the presence
  of highly-annotated genes may cause functionally similar terms to
  end up at the top of the enriched term list (for example, the
  ``response to...'' terms in Table \ref{table_go} contain many
  similar genes), it should not affect the fact that each term is
  independently statistically enriched for \acp{psg}. The \topgo
  method was designed in part to actively correct for this bias, and
  it was for that reason that I included the topGO derived p-values in
  Table \ref{table_go}.}{Gene Annotation Bias}

}

\item{\corr{Low PSG Overlap} Please discuss in more detail why there was so little overlap between
PSGs described in different studies, as shown in Figure 5.4.

\resp{An additional paragraph was added to the relevant section
  further discussing the issue.}{Low PSG Overlap} }

\section{Chapter 6}

\item{\corr{King Wilson Discussion} The discussion of King and
  Wilson's hypothesis should probably be moved to the discussion
  section, as it is more of a distraction in the introduction.

\resp{The aforementioned discussion was secondary to the analysis
  presented in the chapter, so for the (somewhat futile) sake of
  brevity it was removed from the thesis altogether.}{King Wilson
  Discussion} }

\item{\corr{Model Misspecification} Please discuss the potential
  problem of model misspecification when employing models that assume
  different models in different branches. It is entirely possible that
  none of the prespecified models in Figure 6.2 are the 'best', and
  hence any site-by-site models built on top of these may also be
  compromised (see
  http://mbe.oxfordjournals.org/content/28/11/3033.long for a further
  discussion of this). Please consider a wider range of 'alternative'
  models, and compare the goodness-of-fit with the prespecified ones.

\resp{No claim was made that any of the branch models used in this
  chapter would be the ``best'' for any given gene. Rather, the main
  point of using several prespecified models was to sort the set of
  all genes by how well they fit each model, thus identifying the set
  of genes (out of the entire genome-wide set) which are best
  explained by that given pattern. While a detailed analysis of any
  single gene should certainly consider a wider variety of possible
  evolutionary scenarios, when analyzing several thousand genes it
  becomes very difficult to analyze and summarize the results of an
  intensive model comparison. The challenge is not so much in the
  computation, but rather in the interpretation and analysis of many
  model fits across many genes.

  (Informally, I did evaluate a wider range of models by running
  Hyphy's GA branch selection model on the top 10 accelerated genes
  for each of the prespecified models. For the majority of genes the
  best-fit branch model class contained 1 or 2 rates, suggesting that
  a more complicated pattern of evolution rarely better explained the
  data than a one- or two-rate model. However, the limited amount of
  data within these closely-related primate alignments was also
  evident, as the AIC difference between the best-fit and suboptimal
  models was often small, and in some cases the estimated $\omega$
  ratios were extremely high. Given these issues, I decided that the
  partial GA branch analysis did not add enough to the chapter to
  warrant its inclusion.)

  A few paragraphs of new text were added to the chapter explicitly
  discussing the issue of model misspecification.}{Model
  Misspecification}}

\item{\corr{Slightly Deleterious Polymorphism} What might the
  contribution of slightly deleterious polymorphisms be to apparent
  accelerated evolution in humans?

\resp{Text has been added to a new subsection in Chapter 6.}{Slightly
  Deleterious Polymorphism}}

\item{\corr{Considering Polymorphisms} To what extent might the
  results differ if you had multiple genomes per species? What might
  be the contribution of shared polymorphisms in the ancestral species
  in driving apparent accelerations?

\resp{Text has been added to a new subsection in Chapter
  6.}{Considering Polymorphisms}}

\item{\corr{Relevance of dN/dS Differences} Please discuss the
  biological relevance of small (0.01--0.05) differences in
  genome-wide dN/dS (Figure 6.5).

\resp{Text has been added to the new subsection in Chapter
  6.}{Relevance of dN/dS Differences}}

\section{Chapter 7}

\item{Please expand the discussion to include issues such as:
  \begin{itemize}
  \item{The connection between the phylogenetic approaches used here
    and other population-genetic approaches. In particular, please
    comment on why there is an apparent discrepancy between these
    approaches in the proportion of sites under adaptive
    evolution. How does the finding of positive selection relate to
    change in phenotype/fitness, and its magnitude?}
  \item{Please comment on the limitations of the approaches used
    here--the restriction to analysis of coding regions, detection of
    particular types of selection, etc.}
  \item{Given that low-coverage genomes were used in the analysis,
    please comment on whether sequencing errors may have affected the
    results.}
  \end{itemize}
}

\section{General comments}

\item{[Acronyms] Please include a table of abbreviations.

\resp{Done.}{Acronyms}
}

\item{Please fix references (e.g.\ [Yang]).}

\item{Typographical errors:
  \begin{itemize}
  \item{Page 1, line 2. 'evoluion' should be evolution.}
  \item{Page 50, paragraph 4. “This chapter begins with an overview
    OF”.}
  \item{Page 75, paragraph 3. 'identify' should be 'identity'.}
  \item{Page 81, last paragraph. 'twp' should be 'two'.}
  \item{Page 96, paragraph 5, 'clsutered' should be 'clustered'.}
  \item{Page 148, first sentence. Missing beginning of sentence.}
  \item{Page 150, second paragraph. “designed BY all three of us”.}
  \end{itemize}
}
\end{enumerate}
