\begin{acknowledgements}      %this creates the heading for the acknowlegments

First and foremost I would like to thank Nick Goldman, who supervised
my Ph.D. research at the EBI. I arrived with little knowledge and even
less experience; it is a testament to Nick's patience, encouragement,
and unequivocal support that I arrived at the other end some sort of
scientist. Thanks especially for giving me the freedom to make my own
mistakes---and for having the kindness to help mend them.

For enlightening my mind and enriching my spirits at work and beyond,
I have the past, former and current members of the Goldman Group to
thank. One would have some difficulty casting a more insightful,
helpful, and playfully skeptical ensemble. The characters, in order of
appearance: Tim, Ari, Martin, Fabio, Jacky, Stefan, Emeric, Botond,
and Hazel (plus the cameos and guest appearances).

My work and my scientific growth have benefitted greatly from fruitful
discussions and productive collaborations with several others who got
caught in my academic web. My thesis advisory committee (Ewan Birney,
Nick Mundy, and Jan Korbel) provided useful feedback and well-tempered
advice; members of the Ensembl Compara team (Albert Vilella, Javier
Herrero) kindly introduced me to their systems and taught me
invaluable ``farming techniques''; and other collaborators (including
Stephen Montgomery, Aylwyn Scally, Tim O'Connor, Amalio Telenti, David
Aanansen and the members of the Mammalian Genome Project Consortium)
have together taught me innumerable lessons about how to do good work
and play nice with others.

To my family, I owe everything for their love and support over the
past years. I couldn't have dreamt of feeling so close to home while
living thousands of miles away. Sure, Skype helped, but it was a
feeling much more magical than the Internet that kept me close.

Throughout the years I've been blessed with great friends, who
together made sure life in Cambridge never lost its lustre. My adopted
Gates ``family'' embraced me with entertaining dinners, exciting trips
and a new bunch of lifelong friends. Juggling pals Ben and Guy allowed
me a curious peek into the world of May ball entertainment. And my
amazing housemates, the fearless tenants of the pink house on
Coleridge Road---Niko, Lesley, Loizos, Jan, Felix, Tom and Karolina
through the years---have always been there when I most needed to
laugh, commisserate, eat good food, or enjoy yet another alphabetical
party.

Finally, the EBI Predocs \& Friends have been there all along, a core
to my life in so many ways. The gang of 2007 predocs---Michele,
Markus, Adam, Diva, Julia and Judith---have been constant companions
from Heidelberg to Cambridge, as well as reliable scientists, friends,
and travelers on our various acronymic holidays. To all the others, I
am thankful for the never-ending mailing list hijinks, the 24-hour
support network (for computers and for life) and the endless good
cheer. It's a rare bunch that can keep you looking forward to what's
next even after four years. And I would like to extend a very special
thank-you to Anna, who as a friend and so much more has filled this
``thesis year'' of our lives with a much-needed blend of caring,
encouragement and fun.

\end{acknowledgements}
